---
title: "Files formats & Predicting Future Sales"
output:
  html_document: default
  html_notebook: default
---

## Getting data

Download all datasets from  <https://www.kaggle.com/c/competitive-data-science-predict-future-sales/data> into directory `future_sales_data` and uznip it.

## Access to data from spark


```{r}
library(sparklyr)
library(dplyr)
sc <- spark_connect(master = "local")
```

## Parquet and buckets

```{r}
sales_sdf <- spark_read_csv(sc, "sales", "../future_sales_data/sales_train.csv.gz")

sales_sdf %>%
  mutate(dt=to_timestamp(unix_timestamp(date, 'dd.MM.yyyy'))) %>%
  mutate(year=year(dt), month=month(dt)) %>%
  select(-dt) ->
  sales_sdf
```

```{r}
sales_sdf %>%
  group_by(shop_id)
  summarise
```

```{r}
colname <-"item_id"
sales_sdf %>%
  group_by(colname) %>%
  summarize
```


```{r}
colname <- 'shop_id'
sales_sdf %>%
  group_by_(colname) %>%
  summarise %>%
  summarise(
    n(),
    max(!!rlang::sym(colname), na.rm = TRUE),
    min(!!rlang::sym(colname), na.rm = TRUE)
  ) 
```

```{r}
sales_sdf %>%
  spark_write_parquet(
    "../future_sales_data/sales_train.parquet",
    partition_by = c("shop_id", "year", "month"))
```


```{r}
sales_sdf <- spark_read_parquet(sc, "sales", "../future_sales_data/sales_train.parquet/shop_id=0/year=2013/month=1/")
sales_sdf
```

```{r}
sales_sdf <- spark_read_parquet(sc, "sales", "../future_sales_data/sales_train.parquet/shop_id={0, 1}/year=2013/month=1/")
```

### There is no year

```{r}
sales_sdf %>%
 head
```

```{r}
sales_sdf %>%
  mutate(year=2013) %>%
  head
```

### Function to read from multiple sources

```{r}
library(whisker)
```

```{r}
sales_sdf <- read_sales(0, 2000, 1)
sales_sdf
```

```{r}
```

```{r}
source("salesData.R")
sales_sdf <- read_sales(c(0, 1), 2013, 1)
sales_sdf %>%
  group_by(shop_id) %>%
  summarise()
```

```{r}
start_time <- Sys.time()
sales_sdf <- read_sales(0:59, c(2013, 2014, 2015), 1:12)
sales_sdf %>%
  group_by(shop_id) %>%
  summarise %>%
  print
end_time <- Sys.time()
end_time - start_time
```

About 5 mins on mac...

```{r}
sales_sdf %>%
 head
```

### Problem
Could you move this function to a separate file `salesData.R`, import it from there and use it here?


## Project 1: Predict shop's future sales (simplified version)

We want predict next month total sales of a given shop.

Steps to do:

1. Data preparation
2. Train-test splitting   # with the past data we want to predict future data. Test=last month
3. Training simple linear model  # mlib spark
4. Evaluating: implementing MSE & R2 


1.

```{r}
sales_sdf <- spark_read_csv(sc, "sales", "../future_sales_data/sales_train.csv.gz")
```

```{r}
sales_sdf %>%
  head

sales_sdf %>%
  mutate(dt=to_timestamp(unix_timestamp(date, 'dd.MM.yyyy'))) %>%
  mutate(year=year(dt), month=month(dt)) %>%
  select(-dt) ->
  sales_sdf

monthly_sdf <-
sales_sdf %>%
  group_by(year, month,shop_id)%>%
  summarise(total_items=sum(item_cnt_day,na.rm=TRUE)) 

sdf_register(monthly_sdf, "sales_monthly")
```

```{r}
library(DBI)
dbGetQuery(
  sc,
"SELECT *
,  LAG(total_items) OVER (PARTITION BY shop_id ORDER BY year, month) AS prev_total_items
FROM sales_monthly") %>%
  mutate(prediction=ifelse(is.nan(prev_total_items),0,prev_total_items))->
  final_sdf

final_sdf

```

```{r}
monthly_sdf%>%
  group_by(shop_id) %>%
  mutate(prev_total_items=lag(total_items))
```


2.

```{r}
 sales_sdf%>%
  group_by(year,month)%>%
  summarise%>%
  arrange(desc(year), desc(month))

train_sdf<- final_sdf %>%
  filter(!(year==2015 & month==10))

test_sdf<- final_sdf %>%
  filter(year==2015 & month==10)

test_sdf 

```

```{r}
test_sdf %>%
  summarise(mean(total_items, na.rm=TRUE))

```



3.
```{r}

train_sdf%>%
  ml_linear_regression(total_sales ~ shop_id + year + month) ->
  model

model
```

4. 

```{r}
train_sdf %>%
  sdf_predict(model) %>%
  mutate(res=total_sales - prediction)%>%
  summarise(mean(res*res))   #implement mean(log(y+1)-log(y'+1))^2
  print

test_sdf %>%
  sdf_predict(model) %>%
  mutate(res=total_sales - prediction)%>%
  summarise(mean(res*res))
  print
```

```{r}
train_sdf %>%
  filter(year==2015 & month==9) %>%
  mutate(prediction=total_sales)%>%
  select(-total_sales)%>%
  right_join(test_sdf, by=c("shop_id", "year"))
```


